### 2016-02-18 model/overflowmcmc.tex
Mostly updated the MCMC document to reflect current code - missing some details for the d's and for the algorithms for interesting sub-models

\subsection{Transaction arrival process}
We will assume that transactions arrive according to a non-homogenous Poisson process. Let $N(\tau)$ denote the number of transactions that have arrive by time $\tau$, and let $\lambda(\tau)$ denote some intensity function such that $\lambda(\tau) > 0$ for all times $\tau$. Then any collection of non-overlapping increments $\left\{N(\tau_i + \delta_i) - N(\tau_i)\right\}$ where $\delta_i>0$ for all $i$ are independent with distribution
\begin{align*}
N(\tau_i + \delta_i) - N(\tau_i) \sim Poi(\Lambda(\tau_i + \delta_i) - \Lambda(\tau_i))
\end{align*}
where 
\begin{align*}
\Lambda(\tau) = \int_0^\tau \lambda(u)du.
\end{align*}
Suppose the intensity function is a gamma process with shape function $\Psi(\tau)$ and rate parameter $\phi$. In other words
\begin{align*}
\Lambda(\tau_i + \delta_i) - \Lambda(\tau_i) \stackrel{ind}{\sim} G(\Psi(\tau_i + \delta_i) - \Psi(\tau_i), \phi)
\end{align*}
with mean $[\Psi(\tau_i + \delta_i) - \Psi(\tau_i)]/\phi$. Then for block $k=1,2,\dots,n_{b}$ define $\lambda_k = \Lambda(\tau_k) - \Lambda(\tau_{k-1})$, $\eta_k = N(\tau_k) - N(\tau_{k-1})$, and $\psi_k = \Psi(\tau_k) - \Psi(\tau_{k-1})$. Then assume
\begin{align*}
\eta_k | \lambda_{1:n_{b}} &\stackrel{ind}{\sim} Poi(\lambda_k)\\
\lambda_k|\delta_{1:n_{b}} &\stackrel{ind}{\sim} G(\psi_k, \phi).
\end{align*}

In order for the gamma process to be well defined we require $\phi>0$ and $\Psi(\tau)$ to be nondecreasing and right continuous with $\Psi(0)=0$. An easy way to achieve this is to define $\Psi(\tau) = \int_0^\tau\psi(t)dt$ where $\psi(t)$ is continuous and nonnegative. While $\psi(t)$ can be thought of as the intensity of the gamma process it is also intimately related to the intensity of the Poission process. The mean of $N(\tau + \delta) - N(\tau)$ is 
\begin{align*}
E[N(\tau + \delta) - N(\tau)] &= E\{E[N(\tau + \delta) - N(\tau)|\Lambda(\tau + \delta) - \Lambda(\tau)]\} = E[\Lambda(\tau + \delta) - \Lambda(\tau)]\\
 &= [\Psi(\tau + \delta) - \Psi(\delta)]/\phi = \int_{\tau}^{\tau + \delta}\psi(t)dt/\phi.
\end{align*}
In this way, $\psi(t)/\phi$ can be seen as the expected intensity function of the $N(\tau)$ process. The gamma-Poisson structure also allows for a more flexible model on $N(\tau)$ by allowing for overdispersion. In a Poisson model the mean and variance are the same, but in this case they are proportional:
\begin{align*}
V[N(\tau + \delta) - N(\tau)] &= E\{V[N(\tau + \delta) - N(\tau)|\Lambda(\tau + \delta) - \Lambda(\tau)]\} + V\{E[N(\tau + \delta) - N(\tau)|\Lambda(\tau + \delta) - \Lambda(\tau)]\}\\
&= E[\Lambda(\tau + \delta) - \Lambda(\tau)] + V[\Lambda(\tau + \delta) - \Lambda(\tau)]\\
&= [\Psi(\tau + \delta) - \Psi(\tau)](1/\phi + 1/\phi^2) = \frac{\phi + 1}{\phi}E[N(\tau + \delta) - N(\tau)].
\end{align*}
This allows for overdispersion but not underdispersion since $(\phi + 1)/\phi\in[1,\infty)$ for $\phi\ge 0$.

To complete this portion of the model we need to specify $\psi(t)$. A flexible structure that captures many modeling choices is $\psi(t) = \exp(\bm{z}(t)'\bm{\beta})$ where $\bm{z}(t)' = (z_1(t),\dots,z_p(t))$ is a vector of time-varying covariates continuously differentiable in time and $\bm{\beta}'=(\beta_1,\dots,\beta_p)$ is a vector of regression coefficients, for example $\psi(t) = \exp(\beta_1 + t\beta_2)$ in order to capture a simple trend. In that case we have
\begin{align*}
\Psi(\tau+\delta) - \Psi(\tau) &= \int_{\tau}^{\tau + \delta} e^{\bm{z}(t)'\bm{\beta}}dt = \int_\tau^{\tau + \delta} e^{\beta_1 + t\beta_2}dt = \frac{e^{\beta_1 + \tau\beta_2}}{\beta_2}\left(e^{\delta\beta_2} - 1\right)
\intertext{and thus}
\psi_k &= \frac{e^{\beta_1 + \tau_{k-1}\beta_2}}{\beta_2}\left(e^{\delta_k\beta_2} - 1\right).
\end{align*}
\subsection{Transaction data}
Let $i=1,2,\dots,n_d \equiv N_{n_b}$ index transactions and $d_i$ denote the amount of data for transaction $i$ measured in MBs. Then we assume that for $i=1,2,\dots,n_d$
\begin{align*}
d_i \stackrel{iid}{\sim} G(\alpha_d,\beta_d)
\end{align*}
with mean $\alpha_d/\beta_d$. Along with the Poisson process for $N(\tau)$ this implies that the amount of transaction data that has arrived by time $\tau$, $D(\tau)$, is a compound non-homogenous Poisson process conditional on $\Lambda(\tau)$.

\subsection{Overflow and observations of $M_k$ and $D_k$}
Each block has an upper limit for the amount of transaction data it can hold defined by the block miner. Denote this upper limit by $\bar{D}_k$ --- because different miners use different limits and the same miner uses different limits at different times, each block has its own upper limit. Then the number of transactions in blocks $1,2,\dots,k$ combined, $M_k$, could potentially be less than the number of transactions that have arrived by the time the $k$'th block arrives, $N_k$. Specifically:
\begin{align*}
M_k &= \max\left\{m\in\left\{M_{k-1}, M_{k-1} + 1, \dots, N_k\right\}: \sum_{i = M_{k-1} + 1}^m d_i \le \bar{D}_k\right\}
\end{align*}
so that
\begin{align*}
D_k &= \sum_{i = M_{k-1} + 1}^{M_k} d_i
\end{align*}
with the convention that empty summations are defined as zero, i.e. $\sum_{i=5}^4d_i \equiv 0$.

\section{Priors}
The model has the following unknown parameters: $\sigma^2$, $\phi$, $\beta$, $\alpha_d$, $\beta_d$, and $\bar{D}_k$ for $k=1,2,\dots,n_b$. We will suppose that these parameters are mutually independent in the prior, i.e.
\begin{align*}
p(\sigma^2,\phi,\beta,\alpha_d,\beta_d,\bar{D}_1,\dots,\bar{D}_{n_{b}}) = p(\sigma^2)p(\phi)p(\beta)p(\alpha_d)p(\beta_d)p(\bar{D}_1)\hdots p(\bar{D}_{n_{b}}).
\end{align*}

@@

\subsection{Prior on $\phi$}
The conditionally conjugate prior for $\phi$ is a gamma distribution, so we will suppose that $\phi \sim G(a_{\phi}, b_{\phi})$. To choose $a_\phi$ and $b_\phi$ recall that $\phi$ is a dispersion parameter:
\begin{align*}
V[N(\tau)|\phi,\Psi(\tau)] = \frac{\phi + 1}{\phi}E[N(\tau)|\phi,\Psi(\tau)] = \frac{\phi + 1}{\phi}\frac{\Psi(\tau)}{\phi}.
\end{align*}
Setting $\phi = 1$ means that the variance of $N(\tau)$ is twice its mean. Absent strong prior information about $\phi$, we will center our prior on $2$ with a high level of uncertainty. The mean of the gamma distribution is $a_\phi/b_\phi$ with variance $a_\phi/b_\phi^2$, so we will set $b_\phi = 1/100$ to reflect a high degree of uncertainty, and $a_\phi = 1/50$ so that $a_\phi/b_\phi = 2$.

\subsection{Prior on $\bm{\beta}$}
As we will see in the next section, there is no convenient conditionally conjugate form for $\bm{\beta}$, so there is no computational trade-off associated with choosing a prior which more accurately reflects prior knowledge. We will consider $\bm{\beta}$ much like a regression parameter and set
\begin{align*}
\bm{\beta} \sim N(\bm{b}_\beta, \bm{S}_\beta)
\end{align*}
where $\bm{b}_\beta$ is a mean vector and $\bm{S}_\beta$ is a covariance matrix. We will use a weakly informative prior with $\bm{b}_\beta = \bm{0}$ and $\bm{S}_\beta = s^2_\beta\bm{I}$ where $s_\beta = 100$.  [MAYBE SOMETHING WITH FATTER TAILS HERE? MAYBE NOTE THAT NONINFORMATIVE PRIOR IS NOT A GOOD IDEA SINCE WE WANT TO GUARANTEE POSTERIOR PROPRIETY --- SAME FOR $s_\beta$ TOO LARGE (VIVEK'S PAPER)]

\subsection{Priors on $\alpha_d$ and $\beta_d$}
The conditionally conjugate prior for $beta_d$ is a gamma while $\alpha_d$ has no convenient conditionally conjugate form. We will assume that $\beta_d \sim G(a_\beta, b_\beta)$ and $\alpha_d \sim G(a_\alpha, b_\alpha)$. In order to choose $a_\alpha$, $b_\alpha$, $a_\beta$, and $b_\beta$, consider that $E[d_i|\alpha_d,\beta_d] = \alpha_d/\beta_d$. But 
\begin{align*}
E\left[\frac{\alpha_d}{\beta_d}\right] = E[\alpha_d]E\left[\frac{1}{\beta_d}\right] = \frac{a_\alpha}{b_\alpha}\frac{b_\beta}{a_\beta - 1}.
\end{align*}
Similarly
\begin{align*}
V\left[\frac{\alpha_d}{\beta_d}\right] = V[\alpha_d] + V\left[\frac{1}{\beta_d}\right] = \frac{a_\alpha}{b_\alpha^2} + \frac{b_\beta^2}{(a_\beta - 1)^2(a_\beta - 2)}.
\end{align*}
We will choose these parameters so that $E[\alpha_d/\beta_d] \approx 0.01$ MB and $V[\alpha_d/\beta_d]$ is large so that the prior is weakly informative. We will set $a_\beta = 2.1$ to ensure the variance calculated above exists. Then $a_\alpha = 0.1$, $b_\alpha = 1$, and $b_\beta = 1$ yields
\begin{align*}
E\left[\frac{\alpha_d}{\beta_d}\right] = \frac{1}{21} \approx 0.09; && V\left[\frac{\alpha_d}{\beta_d}\right] \approx 10.
\end{align*}

\subsection{Priors on the $\bar{D}_k$'s}
We will put a discrete prior on each of the $\bar{D}_k$'s. Theoretically a miner can put any upper limit she wants on a mined bitcoin, but in practice most miners use one of several possible values. We will use the set of values $\mathcal{D} = \{0.25, 0.35, 0.5, 0.75, 0.9, 0.95, 1\}$ [DOUBLE CHECK THESE --- IN PARTICULAR, ARE THERE OTHER COMMON VALUES? WHAT ABOUT THE ELGIUS MINING POOL'S CAPS?]. For each $\bar{D}_k$ we will use a discrete uniform prior on $\mathcal{D}$ so that $P(\bar{D}_k = j) = 1/7$ for all $j \in \mathcal{D}$.

@@

\subsection{Block arrival times --- $\tau_k$}
The full conditional distribution of the block arrival times is more complicated:
\begin{align*}
p(\bm{\tau}_{1:n_{b}}|...) \propto \exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k+1}\left(\tilde{t}_{ik} - \tau_k\right)^2 - \frac{1}{10}\tau_{n_{b}}\right]\prod_{k=1}^{n_b}\frac{\phi^{\Psi(\tau_k) - \Psi(\tau_{k-1})}}{\Gamma\left[\Psi(\tau_k) - \Psi(\tau_{k-1})\right]}\lambda_k^{\Psi(\tau_k)-\Psi(\tau_{k-1}) - 1}
\end{align*}
where $\tau_0\equiv0$ and $\Gamma(x) = \int_0^\infty t^{x-1}e^{-t}dt$ is the gamma function. Even if $\Psi(\tau)$ is some nice known form, this density is nonstandard because the $\tau_k$'s enter the gamma function. Some ideas for handling this step:
\begin{enumerate}
\item Draw each $\tau_k$ condtional on all the others using random walk Metropolis steps for $k=1,\dots,n_{b}$.
\item Approximate the density of $\tau_k$ sufficiently well and use independent Metropolis steps for $k=1,\dots,n_{b}$.
\item Approximate the density of $\tau_{1:n_{b}}$ sufficiently well and use an independent Metropolis step.
\item Use adaptive rejection sampling (ARS) to draw from the density of $\tau_k$.
\item Use ARMS --- ARS except with a Metropolis correction for densities which are not log concave (look it up).
\end{enumerate}

\subsection{Transaction process dispersion --- $\phi$}
The full conditional density of $\phi$ is
\begin{align*}
p(\phi|...) \propto \prod_{k=1}^{n_{b}}\phi^{\psi_k}e^{\lambda_k\phi}p(\phi) \propto \phi^{\sum_{k=1}^{n_{b}}\psi_k}e^{\phi\sum_{k=1}^{n_{b}}\lambda_k}p(\phi).
\end{align*}
With $\phi \sim G(a_\phi, b_\phi)$ in the prior, this is a gamma density so the step is simple:
\begin{enumerate}
\item Draw $\phi \sim G\left(a_\phi + \sum_{k=1}^{n_{b}}\psi_k, b_{\phi} + \sum_{k=1}^{n_{b}}\lambda_k \right)$.
\end{enumerate}

\subsection{Regression parameters --- $\bm{\beta}$}
The full conditional posterior of $\bm{\beta}$ is closely connected to the functional form of the covariates in time. No matter the functional form, however, the density is complex and a Metropolis step is likely warranted. Recall that since we defined $\psi(t) = e^{\bm{z}(t)'\bm{\beta}}$ we have $\psi_k(\bm{\beta}) = \int_{\tau_{k-1}}^{\tau^k}e^{\bm{z}(t)'\bm{\beta}}dt$ --- in this subsection we write $\psi_k(\bm{\beta})$ to explicitly acknowledge that $\psi_k$ is a function of $\bm{\beta}$. Then the full conditional of $\bm{\beta}$ is
\begin{align*}
p(\bm{\beta}|\dots) \propto p(\bm{\beta})\prod_{k=1}^{n_b}\frac{\phi^{\psi_k(\bm{\beta})}\lambda_k^{\psi_k(\bm{\beta}) - 1}}{\Gamma[\psi_k(\bm{\beta})]}
\end{align*}
where in the prior $\bm{\beta} \sim p(\bm{\beta})$ independent of the other parameters. This density is complex and nonstandard since $\psi_k(\bm{\beta})$ enters the gamma function. A Metropolis step is likely warranted here as well.

\subsection{Transaction arrival times and sizes in MB --- $N(\tau_k)$ and $D_k$}
First consider transaction sizes, $\bm{d}=(d_1,d_2,\dots,d_{n_d})'$. Let $\bm{d}_k$ denote the sizes of the $M_k - M_{k-1}$ transactions that were stored in block $k$ so that $\bm{d} = (\bm{d}_1',\dots,\bm{d}_{n_{b}}')'$. The full conditional distribution of the $\bm{d}_k$'s is
\begin{align*}
p(\bm{d}_{1:n_{b}}|...)\propto \prod_{k=1}^{n_b}\prod_{i=1}^{M_k - M_{k-1}} d_i^{\alpha_d - 1}e^{-d_i\beta_d}1\left(\sum_{i=1}^{M_k - M_{k-1}} d_{ik}= D_k\right)1\left(d_{1k} > \bar{D}_{k-1} - D_{k-1} \mathrm{\ or\ } M_{k-1} = N_{k-1}\right).
\end{align*}
The second indicator function is required because if $M_{k-1} < N_{k-1}$ then $N_{k-1} - M_{k-1}$ transactionss must have overflowed into block $k$. So the amount of data from the first transaction of block $k$, $d_{1k}$, must have been greater than the difference between the amount of data in block $k-1$ and the data cap for block $k-1$. Similarly if $M_{k-1} = N_{k-1}$ then we know there was no overflow into block $k$, so we have no reason to think that $d_{1k}$ has a restricted support. Let $\bm{o}_k = \bm{d}_k/D_k$. Then the $\bm{o}_k$'s have potentially truncated independent Dirichlet distributions. That is
\begin{align*}
p(\bm{o}_{1:n_b}|...)\propto \prod_{k=1}^{n_b} \prod_{i=1}^{M_k - M_{k-1}}o_{ik}^{\alpha_d - 1} 1\left(d_{1k} > [\bar{D}_{k-1} - D_{k-1}]/D_{k} \mathrm{\ or\ } M_{k-1} = N_{k-1}\right).
\end{align*}
So the $\bm{d}$ step to draw from $p(\bm{d}|...)$ is as follows: Let $\bm{1}_n$ denote an $n$-vector of ones. For $k=1,2,\dots,n_b$
\begin{enumerate}
\item If $N_k = M_k$ draw $\bm{o}_{k} \sim Dir(\alpha_d\bm{1}_{M_k - M_{k-1}})$.

>>>>

\subsection{Transaction arrival process}
We will assume that transactions arrive according to a non-homogenous Poisson process. Let $N(\tau)$ denote the number of transactions that have arrive by time $\tau$, and let $\lambda(\tau)$ denote some intensity function such that $\lambda(\tau) > 0$ for all times $\tau$. Then any collection of non-overlapping increments $\left\{N(\tau_i + \delta_i) - N(\tau_i)\right\}$ where $\delta_i>0$ for all $i$ are independent with distribution
\begin{align*}
N(\tau_i + \delta_i) - N(\tau_i) \sim Poi(\Lambda(\tau_i + \delta_i) - \Lambda(\tau_i))
\end{align*}
where 
\begin{align*}
\Lambda(\tau) = \int_0^\tau \lambda(u)du.
\end{align*}
Suppose the intensity function is a gamma process with shape function $\Psi(\tau)$ and rate parameter $\theta$. In other words
\begin{align*}
\Lambda(\tau_i + \delta_i) - \Lambda(\tau_i) \stackrel{ind}{\sim} G(\Psi(\tau_i + \delta_i) - \Psi(\tau_i), \theta)
\end{align*}
with mean $[\Psi(\tau_i + \delta_i) - \Psi(\tau_i)]/\theta$. Then for block $k=1,2,\dots,n_{b}$ define $\lambda_k = \Lambda(\tau_k) - \Lambda(\tau_{k-1})$, $\eta_k = N(\tau_k) - N(\tau_{k-1})$, and $\psi_k = \Psi(\tau_k) - \Psi(\tau_{k-1})$. Then we assume
\begin{align*}
\eta_k | \lambda_{1:n_{b}} &\stackrel{ind}{\sim} Poi(\lambda_k)\\
\lambda_k|\delta_{1:n_{b}} &\stackrel{ind}{\sim} G(\psi_k, \theta).
\end{align*}

In order for the gamma process to be well defined we require $\theta>0$ and $\Psi(\tau)$ to be nondecreasing and right continuous with $\Psi(0)=0$. An easy way to achieve this is to define $\Psi(\tau) = \int_0^\tau\psi(t)dt$ where $\psi(t)$ is continuous and nonnegative. While $\psi(t)$ can be thought of as the intensity of the gamma process it is also intimately related to the intensity of the Poission process. The mean of $N(\tau + \delta) - N(\tau)$ is 
\begin{align*}
E[N(\tau + \delta) - N(\tau)] &= E\{E[N(\tau + \delta) - N(\tau)|\Lambda(\tau + \delta) - \Lambda(\tau)]\} = E[\Lambda(\tau + \delta) - \Lambda(\tau)]\\
 &= [\Psi(\tau + \delta) - \Psi(\delta)]/\theta = \int_{\tau}^{\tau + \delta}\psi(t)dt/\theta.
\end{align*}
In this way, $\psi(t)/\theta$ can be seen as the expected intensity function of the $N(\tau)$ process. The gamma-Poisson structure also allows for a more flexible model on $N(\tau)$ by allowing for overdispersion. In a Poisson model the mean and variance are the same, but in this case they are proportional:
\begin{align*}
V[N(\tau + \delta) - N(\tau)] &= E\{V[N(\tau + \delta) - N(\tau)|\Lambda(\tau + \delta) - \Lambda(\tau)]\} + V\{E[N(\tau + \delta) - N(\tau)|\Lambda(\tau + \delta) - \Lambda(\tau)]\}\\
&= E[\Lambda(\tau + \delta) - \Lambda(\tau)] + V[\Lambda(\tau + \delta) - \Lambda(\tau)]\\
&= [\Psi(\tau + \delta) - \Psi(\tau)](1/\theta + 1/\theta^2) = \frac{\theta + 1}{\theta}E[N(\tau + \delta) - N(\tau)].
\end{align*}
This allows for overdispersion but not underdispersion since $(\theta + 1)/\theta\in[1,\infty)$ for $\theta\ge 0$.

We can represent this part of the model more compactly by integrating out the $\lambda_k$'s. Define $\phi = \theta/(1 + \theta)$. Then $\eta_k$ has a negative binomial distribution:
\begin{align*}
\eta_k|\delta_{1:n_{b}} &\stackrel{ind}{\sim} NB(\psi_k, 1-\phi)
\end{align*}
with pmf
\begin{align*}
P(\eta_k = \eta|\delta_{1:n_{b}}) \frac{\Gamma(\psi_k + \eta)}{\Gamma(\eta + 1)}\phi^{\psi_k}(1-\phi)^{\eta}
\end{align*}
for $\eta=0,1,\dots$ We will use this representation of the model both for simplified exposition and later for cheaper MCMC.

To complete this portion of the model we need to specify $\psi(t)$. A flexible structure that captures many modeling choices is $\psi(t) = \exp(\bm{z}(t)'\bm{\gamma})$ where $\bm{z}(t)' = (z_1(t),\dots,z_p(t))$ is a vector of time-varying covariates continuously differentiable in time and $\bm{\gamma}'=(\gamma_1,\dots,\gamma_p)$ is a vector of regression coefficients, for example $\psi(t) = \exp(\gamma_1 + t\gamma_2)$ in order to capture a simple trend. In that case we have
\begin{align*}
\Psi(\tau+\delta) - \Psi(\tau) &= \int_{\tau}^{\tau + \delta} e^{\bm{z}(t)'\bm{\gamma}}dt = \int_\tau^{\tau + \delta} e^{\gamma_1 + t\gamma_2}dt = \frac{e^{\gamma_1 + \tau\gamma_2}}{\gamma_2}\left(e^{\delta\gamma_2} - 1\right)
\intertext{and thus}
\psi_k &= \frac{e^{\gamma_1 + \tau_{k-1}\gamma_2}}{\gamma_2}\left(e^{\delta_k\gamma_2} - 1\right).
\end{align*}

\subsection{Transaction data}
Let $i=1,2,\dots,n_d \equiv N_{n_b}$ index transactions and $d_i$ denote the amount of data for transaction $i$ measured in MBs. Then we assume that for $i=1,2,\dots,n_d$
\begin{align*}
d_i \stackrel{iid}{\sim} G(\alpha,\beta)
\end{align*}
with mean $\alpha/\beta$. Along with the Poisson process for $N(\tau)$ this implies that the amount of transaction data that has arrived by time $\tau$, $D(\tau)$, is a compound non-homogenous Poisson process conditional on $\Lambda(\tau)$.

\subsection{Overflow and observations of $M_k$ and $D_k$}
Each block has an upper limit for the amount of transaction data it can hold defined by the block miner. Denote this upper limit by $\bar{D}_k$ --- because different miners use different limits and the same miner uses different limits at different times, each block has its own upper limit. Then the number of transactions in blocks $1,2,\dots,k$ combined could potentially be less than the number of transactions that have arrived by the time the $k$'th block arrives. Let $M_k$ denote the number of transactions total in blocks $1,2,\dots,k$. Then possibly $M_k \le N_k$. Specifically:
\begin{align*}
M_k &= \max\left\{m\in\left\{M_{k-1}, M_{k-1} + 1, \dots, N_k\right\}: \sum_{i = M_{k-1} + 1}^m d_i \le \bar{D}_k\right\}
\end{align*}
so that
\begin{align*}
D_k &= \sum_{i = M_{k-1} + 1}^{M_k} d_i
\end{align*}
with the convention that empty summations are defined as zero, i.e. $\sum_{i=5}^4d_i \equiv 0$.

\section{Priors}
The model has the following unknown parameters: $\sigma^2$, $\phi$, $\beta$, $\alpha$, $\beta$, and $\bar{D}_k$ for $k=1,2,\dots,n_b$. We will suppose that these parameters are mutually independent in the prior, i.e.
\begin{align*}
p(\sigma^2,\phi,\beta,\alpha_d,\beta_d,\bar{D}_1,\dots,\bar{D}_{n_{b}}) = p(\sigma^2)p(\phi)p(\beta)p(\alpha_d)p(\beta_d)p(\bar{D}_1)\hdots p(\bar{D}_{n_{b}}).
\end{align*}

@@

\subsection{Prior on $\phi$}
The conditionally conjugate prior for $\phi$ is a beta distribution, so we will suppose that $\phi \sim B(a_{\phi}, b_{\phi})$. To choose $a_\phi$ and $b_\phi$ recall that $\phi$ is related to $\theta$, a dispersion parameter:
\begin{align*}
V[N(\tau)|\theta,\Psi(\tau)] = \frac{\theta + 1}{\theta}E[N(\tau)|\theta,\Psi(\tau)] = \frac{1}{\phi}E[N(\tau)|\theta,\Psi(\tau)] = \frac{1}{\phi}\Psi(\tau)\frac{1-\phi}{\phi}.
\end{align*}
Setting $\phi = 0.5$ means that the variance of $N(\tau)$ is twice its mean. Absent strong prior information about $\phi$, we will center our prior on $2/3$ with a high level of uncertainty. The mean of the beta distribution is $a_\phi/(a_\phi + b_\phi)$ with variance $a_\phi b_\phi/[(a_\phi + b_\phi + 1)(a_\phi + b_\phi)^2]$, so we will set $a_\phi = 0.2$ and $b_\phi = 0.1$ to reflect a high degree of uncertainty.

\subsection{Prior on $\bm{\gamma}$}
As we will see in the next section, there is no convenient conditionally conjugate form for $\bm{\gamma}$, so there is no computational trade-off associated with choosing a prior which more accurately reflects prior knowledge. We will consider $\bm{\gamma}$ much like a regression parameter and set
\begin{align*}
\bm{\beta} \sim N(\bm{g}_\gamma, \bm{S}_\gamma)
\end{align*}
where $\bm{g}_\gamma$ is a mean vector and $\bm{S}_\gamma$ is a covariance matrix. We will use a weakly informative prior with $\bm{g}_\gamma = \bm{0}$ and $\bm{S}_\gamma = s^2_\gamma\bm{I}$ where $s_\gamma = 100$.  [MAYBE SOMETHING WITH FATTER TAILS HERE? MAYBE NOTE THAT NONINFORMATIVE PRIOR IS NOT A GOOD IDEA SINCE WE WANT TO GUARANTEE POSTERIOR PROPRIETY --- SAME FOR $s_\gamma$ TOO LARGE (VIVEK'S PAPER)]

\subsection{Priors on $\alpha$ and $\beta$}
The conditionally conjugate prior for $beta$ is a gamma while $\alpha$ has no convenient conditionally conjugate form. We will assume that $\beta \sim G(a_\beta, b_\beta)$ and $\alpha \sim G(a_\alpha, b_\alpha)$. In order to choose $a_\alpha$, $b_\alpha$, $a_\beta$, and $b_\beta$, consider that $E[d_i|\alpha,\beta] = \alpha/\beta$. But 
\begin{align*}
E\left[\frac{\alpha}{\beta}\right] = E[\alpha]E\left[\frac{1}{\beta}\right] = \frac{a_\alpha}{b_\alpha}\frac{b_\beta}{a_\beta - 1}.
\end{align*}
Similarly
\begin{align*}
V\left[\frac{\alpha}{\beta}\right] = V[\alpha] + V\left[\frac{1}{\beta}\right] = \frac{a_\alpha}{b_\alpha^2} + \frac{b_\beta^2}{(a_\beta - 1)^2(a_\beta - 2)}.
\end{align*}
We will choose these parameters so that $E[\alpha/\beta] \approx 0.01$ MB and $V[\alpha/\beta]$ is large so that the prior is weakly informative. We will set $a_\beta = 2.1$ to ensure the variance calculated above exists. Then $a_\alpha = 0.1$, $b_\alpha = 1$, and $b_\beta = 1$ yields
\begin{align*}
E\left[\frac{\alpha}{\beta}\right] = \frac{1}{21} \approx 0.09; && V\left[\frac{\alpha}{\beta}\right] \approx 10.
\end{align*}

\subsection{Priors on the $\bar{D}_k$'s}
We will put a discrete prior on each of the $\bar{D}_k$'s. Theoretically a miner can put any upper limit she wants on a mined bitcoin, but in practice most miners use one of several possible values. We will use the set of values $\mathcal{D} = \{0.25, 0.35, 0.5, 0.75, 0.935, 0.95, 1\}$. For each $\bar{D}_k$ we will use a discrete uniform prior on $\mathcal{D}$ so that $P(\bar{D}_k = j) = 1/7$ for all $j \in \mathcal{D}$.

@@

\subsection{Block arrival times --- $\tau_k$}
The full conditional distribution of the block arrival times is more complicated:
\begin{align*}
p(\bm{\tau}_{1:n_{b}}|...) \propto & \exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k+1}\left(\tilde{t}_{ik} - \tau_k\right)^2 - \frac{1}{10}\tau_{n_{b}}\right]\\
&\times \prod_{k=1}^{n_b}\Gamma[\psi_k(\tau_{k}, \tau_{k-1}) + \eta_k]\phi^{\psi_k(\tau_k,\tau_{k-1})}1(0\le \tau_1 \le \tau_2 \le \hdots \le \tau_{n_b})
\end{align*}
where $\tau_0\equiv0$ and $\Gamma(x) = \int_0^\infty t^{x-1}e^{-t}dt$ is the gamma function. Here we make explicit that $\psi_k$ is a function of $\tau_k$ and $\tau_{k-1}$. Even if $\Psi(\tau)$ is some nice known form, this density is nonstandard because the $\tau_k$'s enter the gamma function. Some ideas for handling this step:
\begin{enumerate}
\item Draw each $\tau_k$ condtional on all the others using random walk Metropolis steps for $k=1,\dots,n_{b}$.
\item Approximate the density of $\tau_k$ sufficiently well and use independent Metropolis steps for $k=1,\dots,n_{b}$.
\item Approximate the density of $\tau_{1:n_{b}}$ sufficiently well and use an independent Metropolis step.
\item Use adaptive rejection sampling (ARS) to draw from the density of $\tau_k$.
\item Use ARMS --- ARS except with a Metropolis correction for densities which are not log concave (look it up).
\end{enumerate}

Let $\delta_k = \tau_k - \tau_{k-1}$ for $k=1,2,\dots,n_b$. Then the determinant of the Jacobian is $1$ full conditional of $\delta_{1:n_b}$ is
\begin{align*}
p(\bm{\delta}_{1:n_{b}}|...) \propto & \exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k+1}\left(\tilde{t}_{ik} - \sum_{l=1}^k\delta_l\right)^2 - \frac{1}{10}\sum_{k=1}^{n_b}\delta_{k}\right]\\
&\times \prod_{k=1}^{n_b}\Gamma(\psi_k + \eta_k)\phi^{\psi_k}1(0\le \delta_k).
\end{align*}
Finally let $\zeta_k = \log(\delta_k)$ for $k = 1, 2, \dots, n_b$. Then the determinate of the Jacobian is $\prod_{k=1}^{n_b}e^{\zeta_k}$ and the full conditional of $\zeta_{1:n_b}$ is
\begin{align*}
p(\bm{\zeta}_{1:n_{b}}|...) \propto & \exp\left[-\frac{1}{2\sigma^2}\sum_{k=1}^{n_{b}}\sum_{i=1}^{\tilde{n}_k+1}\left(\tilde{t}_{ik} - \sum_{l=1}^ke^{\zeta_l}\right)^2 - \frac{1}{10}\sum_{k=1}^{n_b}e^{\zeta_k} + \sum_{k=1}^{n_b}{\zeta_k}\right] \prod_{k=1}^{n_b}\Gamma(\psi_k + \eta_k)\phi^{\psi_k}.
\end{align*}

For an intial sampler, we will use a random walk Metropolis algorithm on each coordinate with a normal proposal distribution $N(\zeta_k, h_k^2)$. Then the acceptance rate of the Metropolis step for $\zeta_k$ is
\begin{align*}
R_k = \frac{p(\zeta_k^{(prop)}|\dots)}{p(\zeta_k^{(old)}|\dots)} = \frac{p(\bm{\zeta}_{1:n_b}^{(prop)}|\dots)}{p(\zeta_{1:n_b}^{(old)}|\dots)}
\end{align*}
where $\zeta_k^{(old)}$ is the previous value in the chain, $\zeta_k^{(prop)}$ is the proposed value, $\bm{\zeta}_{1:n_b}^{(old)}$ is the previous value of the entire vector $\bm{\zeta}_{1:n_b}$, and $\bm{\zeta}_{1:n_b}^{(prop)}$ is the same except for $\zeta_k^{(prop)}$ in the $k$'th coordinate. So the $\zeta_k$ step is as follows: for $k=1,2,\dots,n_b$
\begin{enumerate}
\item Draw $\zeta_k^{(prop)} \sim N(\zeta_k^{(t)},h_k^2)$.
\item Compute $\delta_{1:n_b}^{(prop)}$, $\tau_{1:n_b}^{(prop)}$, and $\psi_{1:n_b}^{(prop)}$.
\item Compute the acceptance rate $R_k$.
\item Set $\zeta_k^{(t+1)} = \zeta_k^{(prop)}$, $\delta_{1:n_b} = \delta_{1:n_b}^{(prop)}$, $\tau_{1:n_b} = \tau_{1:n_b}^{(prop)}$, and $\psi_{1:n_b} = \psi_{1:n_b}^{(prop)}$ with probability $\min(R_k, 1)$ and $\zeta_k^{(t+1)} = \zeta_k^{(t)}$ otherwise. 
\end{enumerate}
For each $k$ the scale parameter $h_k$ must be chosen by the user, which is practically impossible with a large number of blocks in the analysis. So we automatically tune $h_k$ during the the burn in period as follows [CITATION]:
\begin{enumerate}
\item Run the algorithm above for $H$ iterations.
\item For each $k$, compute the observed acceptance rate of the last $H$ iterations: $r_k\#(\zeta_k^{(t)} = \zeta_k^{(t-1)})/H$.
\item If $r_k$ is larger than some target range $[r_l^*, r_h^*]$ update $\log h_k = \log h_k + c$; if lower update $\log h_k = \log h_k - c$.
\end{enumerate}
In certain simple cases the optimal acceptance rate for a one-dimensional target has been computed to be approximately $0.44$, so we use the range $[0.4, 0.45]$ as a rough guideline [CITATION]. In order to ensure that the Markov chain converges to the target posterior of distribution this adaptive process must asymptotically stop in some sense --- restricting the process to the burn in is a sufficient condition. In practice as the chain converges in can help to start $H$ small and $c$ large so that the $h_k$'s quickly move to reasonable values, then over time increase $H$ and decrease $c$ for more precise fine tuning. We start with $H=50$ and $c=0.1$
